FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-06-18 paper.tex
cite gpu fpx
oceananigans output table
ocean example

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FlowFPX}{FlowFPX}
\newcommand{\GPUFPX}{GPU-FPX}
\newcommand{\FloatTracker}{FloatTracker}
\newcommand{\FT}{\FloatTracker}
\newcommand{\Fp}{Floating-point} % hyphen or no?
\newcommand{\fp}{floating-point} % hyphen or no?
\newcommand{\CSTG}{stack graph}
\newcommand{\CPP}{\code{C++}}
\newcommand{\Dendro}{\textsc{Dendro}}
\newcommand{\urlaccess}[2]{\url{#1}}
\newcommand{\Nan}{\code{NaN}}
\newcommand{\NaN}{\Nan}
\newcommand{\Inf}{\code{Inf}}

@@

We introduce \FlowFPX{}, a toolkit for debugging
\fp{} exceptions in an easy and intuitive way~(\cref{s:flowfpx}):
\begin{itemize}
  \item
    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that uses
    Julia's operator overloading to selectively monitor for exceptions and to
    fuzz code for vulnerabilities.
  \item
    To visualize results, \FT{} adapts coalesced stack-trace graphs 
    (CSTGs, or \CSTG{}s)~\cite{hmbdrg-cse-2014} for a whole-program view on the generation,
    propagation, and death of exceptional values.
  \item \GPUFPX{}~[CITE] provides fine-grained insights for GPU kernels.
\end{itemize}
%
Together, the \FlowFPX{} tools have helped debug exceptions in a variety
of applications: from ocean simulations to heat models~(\cref{s:casestudies}).
That said, there is much more to explore.
\FlowFPX{} is a growing project and we welcome feedback from domain experts
on how to best suit their workflow.

@@

@@

% * what about GPUs? companion project at Utah GPU-FPX
% * architecture diagram, CPU + GPU, show Ocean source code
% * replace CPU() with GPU() in code, get output like below & attached
% 
% \begin{verbatim}
% using Oceananigans
% grid = RectilinearGrid(GPU(), size=(128, 128), x=(0, 2π), y=(0, 2π), topology=(Periodic, Periodic, Flat))
% model = NonhydrostaticModel(; grid, advection=WENO())
% ϵ(x, y, z) = 2rand() - 1
% set!(model, u=ϵ, v=ϵ)
% simulation = Simulation(model; Δt=0.01, stop_time=4)
% run!(simulation)
% \end{verbatim}
% 
% ... xinyi ran gpu-fpx on this code ... on the binary it created?
% 
% 
% - - -
% 
% 
% #GPU-FPX LOC-EXCEP INFO: in kernel
% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
% DIV0 found @ ./float.jl:335 [FP64]
% #GPU-FPX LOC-EXCEP INFO: in kernel
% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
% NaN found @ ./float.jl:335 [FP64]
% #GPU-FPX LOC-EXCEP INFO: in kernel
% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
% NaN found @ ./float.jl:335 [FP32]
% #GPU-FPX LOC-EXCEP INFO: in kernel
% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
% DIV0 found @ ./int.jl:261 [FP64]
% #GPU-FPX LOC-EXCEP INFO: in kernel
% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
% NaN found @ ./int.jl:261 [FP64]
% #GPU-FPX LOC-EXCEP INFO: in kernel
% [_Z27julia_broadcast_kernel_716715CuKernelContext13CuDeviceArrayI7ComplexI7Float64ELi3ELi1EE11BroadcastedI12CuArrayStyleILi3EE5TupleI5OneToI5Int64ES6_IS7_ES6_IS7_EE2__S5_IS3_IS4_ILi3EEvS8_S5_I8ExtrudedIS0_IS1_IS2_ELi3ELi1EES5_I4BoolS10_S10_ES5_IS7_S7_S7_EEEES3_IS4_ILi3EEvS8_S5_IS3_IS4_ILi3EEvS8_S5_IS9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EES9_IS0_IS2_Li3ELi1EES5_IS10_S10_S10_ES5_IS7_S7_S7_EEEES7_EEEES7_],
% INF found @ ./int.jl:261 [FP64]
% 
% 
% \begin{verbatim}
% --- GPU-FPX Report ---
% 
% --- FP64 Operations ---
% Total NaN found: 2
% Total INF found: 1
% Total underflow (subnormal): 0
% Total division by zero: 2
% --- FP32 Operations ---
% Total NaN found: 1
% Total INF found: 0
% Total underflow (subnormal): 0
% Total division by zero: 0
% --- Other Stats ---
% Kernels: 21
% \end{verbatim}

>>>>

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FlowFPX}{FlowFPX}
\newcommand{\GPUFPX}{GPU-FPX}
\newcommand{\FloatTracker}{FloatTracker}
\newcommand{\FT}{\FloatTracker}
\newcommand{\Fp}{Floating-point} % hyphen or no?
\newcommand{\fp}{floating-point} % hyphen or no?
\newcommand{\CSTG}{stack graph}
\newcommand{\CPP}{\code{C++}}
\newcommand{\Dendro}{\textsc{Dendro}}
\newcommand{\urlaccess}[2]{\url{#1}}
\newcommand{\Nan}{\code{NaN}}
\newcommand{\NaN}{\Nan}
\newcommand{\Inf}{\code{Inf}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zerocode}[1]{\zerowidth{\code{#1}}}

@@

We introduce \FlowFPX{}, a toolkit for debugging
\fp{} exceptions in an easy and intuitive way~(\cref{s:flowfpx}):
\begin{itemize}
  \item
    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that uses
    Julia's operator overloading to selectively monitor for exceptions and to
    fuzz code for vulnerabilities.
  \item
    To visualize results, \FT{} adapts coalesced stack-trace graphs 
    (CSTGs, or \CSTG{}s)~\cite{hmbdrg-cse-2014} for a whole-program view on the generation,
    propagation, and death of exceptional values.
  \item A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
\end{itemize}
%
Together, the \FlowFPX{} tools have helped debug exceptions in a variety
of applications: from ocean simulations to heat models~(\cref{s:casestudies}).
That said, there is much more to explore.
\FlowFPX{} is a growing project and we welcome feedback from domain experts
on how to best suit their workflow.

@@

% * what about GPUs? companion project at Utah GPU-FPX
% * architecture diagram, CPU + GPU, show Ocean source code
% * replace CPU() with GPU() in code, get output like below & attached

@@

Oceananigans~\cite{OceananigansJOSS} is simulation package for incompressible
fluid dynamics that can generate code for Nvidia GPUs.
For example, the following program (from the project readme) simulates turbulance:

\begin{lstlisting}[language = Julia]
using Oceananigans
grid = RectilinearGrid(GPU(),
  size=(128, 128), x=(0, 2π), y=(0, 2π),
  topology=(Periodic, Periodic, Flat))
model = NonhydrostaticModel(; grid,
  advection=WENO())
ϵ(x, y, z) = 2rand() - 1
set!(model, u=ϵ, v=ϵ)
simulation = Simulation(model;
  Δt=0.01, stop_time=4)
run!(simulation)
\end{lstlisting}

\GPUFPX{} provides detailed feedback on this program.
The output in~\cref{f:gpufpx} shows that 21 kernels
appear and generate six floating-point exceptions.
There are three \NaN{}s, one \Inf{}, and two division
by zero errors.
The report is a starting point for further investigation
of the reliability of the example.
%% TODO exactly how?

\begin{figure}[t]\centering
  \begin{tabular}[t]{ll}
    \begin{tabular}[t]{lr}
      \zerocode{-{}- FP64 Operations -{}-} \\
      \code{Total NaN:} & \code{2} \\
      \code{Total INF:} & \code{1} \\
      \code{Total subnormal:} & \code{0} \\
      \code{Total div0:} & \code{2} \\
    \end{tabular}
    &
    \begin{tabular}[t]{lr}
      \zerocode{-{}- FP32 Operations -{}-} \\
      \code{Total NaN:} & \code{1} \\
      \code{Total INF:} & \code{0} \\
      \code{Total subnormal:} & \code{0} \\
      \code{Total div0:} & \code{0} \\
    \end{tabular}
  \end{tabular}
  \begin{tabular}{lr}
    \zerocode{-{}- Other Stats -{}-} \\
    \code{Kernels:} & \code{21}
  \end{tabular}

  \caption{Example \GPUFPX{} output}
  \label{f:gpufpx}
\end{figure}
